Bayesian Sparsification of Deep $\cplx$-valued networks

### 2020-03-16 paper/paper.tex
manually incorporated edits from the icml branch

The goal of this section to study the compression-performance trade-off of the proposed Bayesian
sparsification methods for $\cplx$-valued networks. To this end we trained each network in
three successive stages in every experiment: ``dense'' $\to$ ``sparsify'' $\to$ ``fine-tune''.
Model parameters and the optimizer are copied between stages, except before the final stage
optimizer's internal state is reset.

@@

% on compresssion
Between \textit{``sparsify''} and \textit{``fine-tune''} stages we compute masks of non-zero
weights in each layer based on the relevance scores $\alpha$ (sec.~\ref{sub:dropout}). Since
the $q_\theta$ factorizes into univariate distributions, a $\cplx$ or $\real$ parameter is
considered non-zero if $\log \alpha \leq \tau$ for $
  \alpha = \tfrac{\sigma^2}{\lvert\mu\rvert^2}
$, or zero otherwise. Figure~\ref{fig:hist__and__threshold__tradeoff} (top) displays the
general shape of the distribution of $\log \alpha$ on the Fashion-MNIST dataset in
sec.~\ref{sub:mnist_like_datasets}. There always appears to be a mode in the moderately
negative range, a valley through zero between the peaks of the negative mode and the next
positive one. The relative mass of each mode changes with $C$ and the higher the coefficient
is, the more mass is shifted to the right.
%
The bottom graph on figure~\ref{fig:hist__and__threshold__tradeoff} is constructed
for $C \in \{\tfrac1{20}, \frac1{200}\}$ before running the ``fine-tune'' stage. From
\eqref{eq:elbo_with_coef}, the relative positions of the compression curves on this
plot, and the results of experiments below it can be concluded that $C$ has much more
substantial impact on the sparsity and performance, than the choice of $\tau$. Furthermore,
for $\tau > 0$ the performance quickly saturates while compression creeps downwards.

\begin{figure}[ht]
  \centering
  \begin{subfigure}[b]{0.95\textwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\linewidth,draft]{../assets/log-alpha__sum__fashion-mnist__20200131-205926.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{0.95\textwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\linewidth]{../assets/figure__musicnet__threshold.pdf}
  \end{subfigure}
  \caption{%
    (\textit{top}) Smoothed density of $\log\alpha$ for some convolutional and dense layer
    from a model in sec.~\ref{sub:mnist_like_datasets};
    (\textit{bottom}) Test split performance and compression rate for a $\cplx$-valued
    net on MusicNet.
  }
  \label{fig:hist__and__threshold__tradeoff}
\end{figure}

% on chosing the threshold
The threshold $\tau$ is picked so, that the remaining \textit{``non-zero''} parameters
are within $\delta$ relative tolerance of their mean with high probability under the
approximate posterior $q_\theta(W)$. This ensures that stochastic parameters are sufficiently
concentrated around $\mu$, which are used as stand-in values for network weights $W$,
when it runs in inference mode. Since $W$ is $\real$- or a proper $\cplx$-Gaussian, the
random variable $
  \tfrac{k \lvert W - \mu \rvert^2}
        {\alpha \lvert \mu \rvert^2}
$ is $\chi^2_k$ distributed with $k=1$ ($\real$) or $2$ ($\cplx$), whence
\begin{equation} \label{eq:prob_relevant}
  \Pr\bigl(
    \lvert w - \mu \rvert \leq \delta \lvert \mu \rvert
  \bigr)
    % = \Pr\biggl(
    %   \frac{k \lvert w - \mu \rvert^2}{\alpha \lvert \mu \rvert^2}
    %     \leq \frac{\delta^2 k}{\alpha}
    % \biggr)
    = \chi^2_k\biggl(\frac{\delta^2 k}{\alpha}\biggr)
  \,.
\end{equation}
For a generous $\delta = 50\%$ tolerance all values of $\log \alpha$ below $-2.5$ yield
at least $90\%$ chance of a non-zero $\real$/$\cplx$ parameter. We pick $\tau = -\tfrac12$
to exclude noisy parameters and encourage higher sparsity, at the same time being aware
that $q_\theta$ is merely an approximation. In comparison, $\tau = 3$ is commonly used as
the threshold \citep{molchanov_variational_2017,kingma_variational_2015}.

@@

% perfromance metrics
We use accuracy as the performance metric for multi-class image classification on MNIST-like
and CIFAR10 datasets. Similarly to \citet{trabelsi_deep_2017} performance in experiments with
MusicNet is measured with (pooled) average precision: predicted scores and target binary labels
for each output are pooled, and average precision is computed based on the precision-recall
curve as if the task was single-output. The evaluation is done after ``dense'' stage, before
the beginning and upon termination of the ``fine-tune'' stage.

% write a note on the incompressible parameters (bias, batch norm etc.)
The compression rate is computed based on the number of floating point values needed to
specify the model and equals $
  \tfrac{n_\mathtt{par}}{n_\mathtt{par} - n_\mathtt{zer}}
$, where $n_\mathtt{zer}$ is the number of explicit zeros after sparsification, $n_\mathtt{par}$
is the total number of floats. Thus, each parameter is counted once in a $\real$-valued
network, but twice in a $\cplx$-network, being a pair of real-imaginary floats. Each model
has incompressible parameters, such as layers' biases, shift and scaling from $\real$-and
$\cplx$-batch normalization layers, if present.

@@

We conduct a moderately sized experiment on MNIST and similar $28\times 28$ greyscale image
datasets to study the compression-performance trade-off of the proposed $\cplx$-variational
dropout on simple architectures.
% dataset size
We deliberately use a {fixed} random subset of $10k$ images from the train split of each
dataset to fit the networks, and gauge the final performance on the usual test split.
Limiting the size was done primarily to test the regularization strength of compression,
and to make bulk experimentation faster.

% model architectures
% discussion of TwoLayerDenseModel (TLDM) and SimpleConvModel (SCM)
The $\real$/$\cplx$ networks, studied in this experiment, have been chosen for the purpose 
of illustrating the compression and understanding the effects of experiment parameters.
\textit{SimpleConvModel} (SCM) is a $2d$ ReLU convolutional network with average pooling
($2\times 2$) after non-linearities with two convolutions (kernel $5\times 5$, stride $1$)
with filters $n_1 \to n_2$ followed by dense layers $n_3 \to n_\mathtt{out}$. The second
network, \textit{TwoLayerDenseModel} (TLDM), is a wide dense network $n_1 \to n_\mathtt{out}$.
For SCM model $n_1, n_2, n_3$ are $20, 50, 500$, respectively, and for TLDM model $
  n_1 = 4096
$ regardless of their type ($\real$ or $\cplx$), although \citet{monning_evaluation_2018}
argue that this renders models incomparable.

@@

% first experiment: 40-75-40 with fast scheduler
In this experiment stages last for $40$, $75$ and $40$ epochs, respectively, and the base
learning rate of ${10}^{-3}$ is reduced after the $10$-th epoch to ${10}^{-4}$.
%
Plots~\ref{fig:mnist-like__trade-off__kmnist} and \ref{fig:mnist-like__trade-off__emnist}
depict samples from the performance-compression curve on KMNIST and EMNIST-Letters datasets
for the studied models and VD compression method (sec.~\ref{ssub:variational_dropout}).
Besides that we tried \texttt{raw} or \texttt{fft} input features, varied $
  C \in \{\tfrac3{2^{k+1}}\colon k=1..16\}
$, and repeated the experiment $5$ times.

We make a couple of observations based on these results. First, Fourier features appear to
perform on par with raw colour channel data in the convolutional networks across most
compression rates, but require high compression rate to make wide shallow dense network work
as well. The upward trend in manifests itself in MNIST and Fashion MNIST experiments as well
(see appendix~\ref{sub:other_mnist_like_datasets}). Finally, overall $\cplx$ variational
dropout methods seems to compress network adequately well, without much loss in performance.

% a couple of plots from MNIST and Fashion-MNIST
\begin{figure}[!h]
  \centering
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__fashionmnist__fft__-0.5.pdf}
    \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__kmnist__fft__-0.5.pdf}
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/VD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__fashionmnist__raw__-0.5.pdf}
    \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__kmnist__raw__-0.5.pdf}
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/VD__kmnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy trade-off on KMNIST dataset.
    (\textit{top}) \texttt{fft} features, (\textit{bottom}) \texttt{raw} features.
  }
  \label{fig:mnist-like__trade-off__kmnist}
\end{figure}

\begin{figure}[!h]
  \centering
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__mnist__fft__-0.5.pdf}
    \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__emnist_letters__fft__-0.5.pdf}
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/ARD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__mnist__raw__-0.5.pdf}
    \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__emnist_letters__raw__-0.5.pdf}
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/VD__emnist_letters__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy trade-off on the letters part of the Extended MNIST dataset.
    (\textit{top}) \texttt{fft} features, (\textit{bottom}) \texttt{raw} features.
  }
  \label{fig:mnist-like__trade-off__emnist}
\end{figure}

@@

% mention the same experimental setup as in MNIST
Experiments on CIFAR10 dataset are set up similarly to MNIST-like experiment, except
that we study VGG16 model \citep{simonyan_very_2015} and its $\cplx$-counterpart and
replicate each experiment only twice. Each network is trained stagewise for $20$, $40$,
and $20$ epochs on the full training split of $50k$ colour $32\times32$ images. We use
only raw colour features (\texttt{raw}). We apply random crop and horizontal flips as
data augmentation and use a learning rate scheduler in SGD, which scales the base learning
rate after $5$, $10$, and $20$-th epoch by $\tfrac1{10}$, $\tfrac1{20}$ and $\tfrac1{100}$,
respectively.

% describe the VGG16 model and its cplx-valued modification
The $\cplx$-VGG16 model is architecturally identical to VGG16, except for all layers being
replaced by their $\cplx$-valued versions. To maintain comparability, we halve the filter
sizes of $\cplx$ model \citep{monning_evaluation_2018}.

@@

% compare compressed cplx-vgg against real-vgg
\begin{figure}[!h]
  \centering
  \begin{subfigure}[b]{0.9\textwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\linewidth]{../assets/figure__cifar__trade-off/legacy__augmentedcifar10__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    Compression-accuracy frontier for $\real$ and $\tfrac12 \cplx$ versions of VGG16 on CIFAR10.
  }
  \label{fig:figure__cifar10__trade-off}
\end{figure}

\begin{figure}[!h]
  \centering
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\linewidth]{../assets/figure__musicnet__trade-off/paper__musicnetram__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    Performance-compression trade-off using \textbf{\color{blue} VD} and \textbf{\color{orange} ARD}.
    The \textbf{\color{pink} version} of the network with kernel $3$ in the first convolution is
    compressed with VD only.
  }
  \label{fig:musicnet__trade-off}
\end{figure}

@@

% describe musicnet as in Trabelsi et al. (2017) also get some hints from thickstun_learning_2017
MusicNet \citep{thickstun_learning_2017} is an annotated audio dataset, consisting of $330$
compositions from the public domain. Folowing \citet{trabelsi_deep_2017}, we use this
dataset for automatic music transcription task and study the $1d$ VGG-like $\cplx$-convent
proposed therein. Similarly to this study, we downsample the audio from 44.1kHz to 11kHz,
retaining only $84$ out of $128$ labels, and hold-out the same validation and test splits.
During each training epoch we sample $1000$ mini-batches of the remaining $321$ compositions,
each mini-batch comprised of a random full window of $4096$ samples picked from each waveform
and transformed from $\real$ to $\cplx$ with $1d$ FFT. Annotation labels are taken at the
mid-point of the window.

Preliminary experiments with the model without sparsification have shown that early stopping
almost always terminates the training process within the first $10-20$ epochs, and that the
validation performance peakeing at about $10-15$ epochs, much sooner than $200$ epochs used by
\citet{trabelsi_deep_2017}. Thus we adopt shorter lasting stages: $12$, $32$ and $50$ epochs,
which essentially turns the ``dense'' stage into pretraining step. Early stopping, tracking
the validation split performance, is used only during the ``fine-tune'' stage. To keep the
learning rate schedule synchronised with such abbreviated training, we allow the base rate
of $10^{-3}$ to be scaled after $5$, $10$, and $20$-th epoch by $\tfrac1{10}$, $\tfrac1{20}$
and $\tfrac1{100}$, respectively. Other key deviations from the set-up used by \citet{trabelsi_deep_2017}
include: shifting lower frequencies to the centre of the window to maintain locality after
applying FFT (sec.~\ref{sub:mnist_like_datasets}), and clipping $\ell_2$ norm of parameters'
gradients to $0.05$.

In each experiment we varied the method and the value $C$, while keeping $\tau$ at $-\tfrac12$.
Each full experiment is replicated five times to take into account random effects from network
initialization before ``dense'', randomness of the mini-batch sequence during SGD, and
stochasticity of intermediate layers on forward pass at ``sparsify'' stage.

Figure~\ref{fig:musicnet__trade-off} shows the resulting performance-compression frontier
for the deep network of \citet{trabelsi_deep_2017} and its version, in which we purposefully
halve the spatial size of the first $1d$ convolution kernel from $6$ to $3$ (denoted by
suffix \textit{k}$3$). The motivation is to test if the performance handicap introduced be
the forced compression of the most upstream layer can be alleviated through non-uniform
compression, induced by variational dropout. We test only VD (sec.~\ref{ssub:variational_dropout})
method in this sub-experiment, since prior results have not demonstrated significant difference
in performance-compression trade-off between VD and ARD (sec.~\ref{ssub:empirical_bayes}).

Shaded horizontal bands represent the min-max spread of the performance of the uncompressed
network upon termination of ``dense'' stage. Each point represents a precision-compression
value attained by a network after ``fine-tune'' stage. The end of the tail of each point
represents the performance gain or loss from the ``fine-tune'' stage itself.
%
For compression levels lower than $50$, the final stage overfits the network, since
it sill remains overparameterized. For rates higher than $\approx 100$ the ``fine-tune''
stage serves its purpose, essentially, due regularization effects if high sparsity. There
is evidence in favour of achievability of the SOTA average precision of \citet{trabelsi_deep_2017}
within the $50$-$100$ compression region with $\cplx$-valued variational dropout with ARD
prior (sec.~\ref{ssub:empirical_bayes}).

@@

Despite having higher arithmetic complexity, complex-valued models provide built-in
joint amplitude-phase transformations and regularization \citep{hirose_complex-valued_2009}
and enjoy superior performance on naturally $\cplx$-valued datasets \citep{tarver_design_2019},
that occur in ultra high-frequency radio communications and audio signal processing. In this
study we address the issue of excessive the arithmetic complexity of complex-valued networks
by extending the existing Bayesian compression techniques by \citet{kingma_variational_2015},
\citet{molchanov_variational_2017} and \citet{kharitonov_variational_2018} to $\cplx$-valued
model parameters. To this end we have conducted a large scale numerical study of such networks
to assess the performance-compression rate trade-off and have achieved performance, comparable
to SOTA on the MusicNet dataset.

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=\linewidth]{../notebooks/assets/grad_log.pdf}
  \caption{$\tfrac{d f(\alpha)}{d \log{\alpha}}$ of the fit proposed by
  \citet{molchanov_variational_2017}, MC estimate of \eqref{eq:improper-kl-div-real},
  and the exact derivative using \eqref{eq:real-kl-div-deriv-log}.}
  \label{fig:molchanov-derivative-replica}
\end{figure}

>>>>

The goal of this section to study the compression-performance trade-off of the proposed Bayesian
sparsification methods for $\cplx$-valued networks. To this end we trained each network in
three successive stages in every experiment: ``dense'' $\to$ ``sparsify'' $\to$ ``fine-tune''.

@@

% on compresssion
Between \textit{``sparsify''} and \textit{``fine-tune''} stages we compute masks of non-zero
weights in each layer based on the relevance scores $\alpha$ (sec.~\ref{sub:dropout}). Since
the $q_\theta$ factorizes into univariate distributions, a $\cplx$ or $\real$ parameter is
considered non-zero if $\log \alpha \leq \tau$ for $
  \alpha = \tfrac{\sigma^2}{\lvert\mu\rvert^2}
$, or zero otherwise. Figure~\ref{fig:hist__and__threshold__tradeoff} is constructed
for $C \in \{\tfrac1{20}, \frac1{200}\}$ before running the ``fine-tune'' stage. From
\eqref{eq:elbo_with_coef}, the relative positions of the compression curves on this
plot, and the results of experiments below it can be concluded that $C$ has much more
substantial impact on the sparsity and performance, than the choice of $\tau$. Furthermore,
for $\tau > 0$ the performance quickly saturates while compression creeps downwards.

\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{0.95\textwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{../assets/figure__musicnet__threshold.pdf}
  \end{subfigure}
  \caption{%
    Test split performance and compression rate for a $\cplx$-valued net on MusicNet.
  }
  \label{fig:hist__and__threshold__tradeoff}
\end{figure}

% on chosing the threshold
The threshold $\tau$ is picked so, that the remaining \textit{``non-zero''} parameters
are within $\delta$ relative tolerance of their mean with high probability under the
approximate posterior $q_\theta(W)$. This ensures that stochastic parameters are sufficiently
concentrated around $\mu$, which are used as stand-in values for network weights $W$,
when it runs in inference mode. Since $W$ is $\real$- or a proper $\cplx$-Gaussian, the
random variable $
  \tfrac{k \lvert W - \mu \rvert^2}
        {\alpha \lvert \mu \rvert^2}
$ is $\chi^2_k$ distributed with $k=1$ ($\real$) or $2$ ($\cplx$).
%
For a generous $\delta = 50\%$ tolerance all values of $\log \alpha$ below $-2.5$ yield
at least $90\%$ chance of a non-zero $\real$/$\cplx$ parameter. We pick $\tau = -\tfrac12$
to exclude noisy parameters and encourage higher sparsity, at the same time being aware
that $q_\theta$ is merely an approximation. In comparison, $\tau = 3$ is commonly used as
the threshold \citep{molchanov_variational_2017,kingma_variational_2015}.

@@

% perfromance metrics
We use accuracy as the performance metric for multi-class image classification on MNIST-like
and CIFAR10 datasets. Similarly to \citet{trabelsi_deep_2017} performance in experiments with
MusicNet is measured with (pooled) average precision. The evaluation is done after ``dense''
stage, before the beginning and upon termination of the ``fine-tune'' stage.

% write a note on the incompressible parameters (bias, batch norm etc.)
The compression rate is computed based on the number of floating point values needed to
store the model and equals $
  \tfrac{n_\mathtt{par}}{n_\mathtt{par} - n_\mathtt{zer}}
$, where $n_\mathtt{zer}$ is the number of explicit zeros, $n_\mathtt{par}$ is the total
number of floats. Thus each parameter is counted once in a $\real$-valued network, but twice
in a $\cplx$-network. Each model has compression limit, set by layers' biases, shift
and scaling from $\real$-and $\cplx$-batch normalization.

@@

We conduct a moderately sized experiment on MNIST and similar $28\times 28$ greyscale image
datasets to study the compression-performance trade-off of the proposed $\cplx$-variational
dropout.
% dataset size
We deliberately use a {fixed} random subset of $10k$ images from the train split of each
dataset to fit the networks and gauge the final performance on the usual test split.
Limiting the size was done primarily to test the regularization strength of compression.

% model architectures
% discussion of TwoLayerDenseModel (TLDM) and SimpleConvModel (SCM)
The $\real$/$\cplx$ networks have been chosen for the purpose of illustrating the compression
and understanding the effects of experiment parameters. \textit{SimpleConvModel} (SCM) is a $2d$
ReLU convolutional network with average pooling ($2\times 2$) after non-linearities with
two convolutions (kernel $5\times 5$, stride $1$) with filters $n_1 \to n_2$ followed by
dense layers $n_3 \to n_\mathtt{out}$. The second network, \textit{TwoLayerDenseModel} (TLDM),
is a wide dense network $n_1 \to n_\mathtt{out}$. For SCM model $n_1, n_2, n_3$ are $20, 50, 500$,
respectively, and for TLDM model $n_1 = 4096$ regardless of their type ($\real$ or $\cplx$),
although \citet{monning_evaluation_2018} argue that this renders models incomparable.

@@

% first experiment: 40-75-40 with fast scheduler
In this experiment stages last for $40$, $75$ and $40$ epochs, respectively, and the base
learning rate of ${10}^{-3}$ is reduced after the $10$-th epoch to ${10}^{-4}$.
%
Plots~\ref{fig:mnist-like__trade-off} depict samples from the performance-compression curve
on KMNIST and EMNIST-Letters datasets for the studied models and VD compression method
(sec.~\ref{ssub:variational_dropout}). Besides that we try \texttt{raw} or \texttt{fft}
input features, vary $
  C \in \{\tfrac3{2^{k+1}}\colon k=1..16\}
$, and repeat the experiment $5$ times.

Shaded horizontal bands reflect the spread of the performance of the uncompressed network
after ``dense'' stage. Each point represents a precision-compression value attained by the
network after ``fine-tune'' stage. The end of each point's tail represents the performance
gain or loss from running the ``fine-tune'' stage.
%
We make a couple of observations based on the results of these plots. First, \texttt{fft}
features perform on par with raw colour channel data in the convolutional networks across most
compression rates, but require high compression rate to make wide shallow dense network work
as well. The upward trend manifests itself in MNIST and Fashion MNIST experiments as well
(see appendix~\ref{sub:other_mnist_like_datasets}). Finally, overall $\cplx$ variational
dropout methods seem to compress networks adequately well without much loss in performance.

% a couple of plots from MNIST and Fashion-MNIST
\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__fashionmnist__fft__-0.5.pdf}
    \includegraphics[width=0.8\columnwidth]{../assets/figure__mnist-like__trade-off/legacy__VD__kmnist__fft__-0.5.pdf}
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/VD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__fashionmnist__raw__-0.5.pdf}
    \includegraphics[width=0.8\columnwidth]{../assets/figure__mnist-like__trade-off/legacy__VD__kmnist__raw__-0.5.pdf}
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/VD__kmnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__mnist__fft__-0.5.pdf}
    \includegraphics[width=0.8\columnwidth]{../assets/figure__mnist-like__trade-off/legacy__VD__emnist_letters__fft__-0.5.pdf}
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/ARD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__mnist__raw__-0.5.pdf}
    \includegraphics[width=0.8\columnwidth]{../assets/figure__mnist-like__trade-off/legacy__VD__emnist_letters__raw__-0.5.pdf}
    % \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/VD__emnist_letters__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy trade-off on KMNIST and the letters of the Extended MNIST datasets.
    (\textit{top}) \texttt{fft} features, (\textit{bottom}) \texttt{raw} features.
  }
  \label{fig:mnist-like__trade-off}
\end{figure}

@@

% mention the same experimental setup as in MNIST
Experiments on CIFAR10 dataset are set up similarly to MNIST-like experiment, except
that we study VGG16 model \citep{simonyan_very_2015} and its $\cplx$-counterpart and
replicate each experiment only twice.
%
% describe the VGG16 model and its cplx-valued modification
The $\cplx$-VGG16 model is architecturally identical to VGG16, except for all layers being
replaced by their $\cplx$-valued versions. To maintain comparability, we halve the filter
sizes of $\cplx$ model \citep{monning_evaluation_2018}.
%
Each network is trained stagewise for $20$, $40$,
and $20$ epochs on the full training split of $50k$ colour $32\times32$ images. We use
only raw colour features (\texttt{raw}). We apply random crop and horizontal flips as
data augmentation and use a learning rate scheduler in SGD, which scales the base learning
rate after $5$, $10$, and $20$-th epoch by $\tfrac1{10}$, $\tfrac1{20}$ and $\tfrac1{100}$,
respectively.

@@

% compare compressed cplx-vgg against real-vgg
\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{0.9\textwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{../assets/figure__cifar__trade-off/legacy__augmentedcifar10__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    Compression-accuracy frontier for $\real$ and $\tfrac12 \cplx$ versions of VGG16 on CIFAR10.
  }
  \label{fig:figure__cifar10__trade-off}
\end{figure}

@@

% describe musicnet as in Trabelsi et al. (2017) also get some hints from thickstun_learning_2017
MusicNet \citep{thickstun_learning_2017} is an annotated audio dataset, consisting of $330$
musical compositions. Folowing \citet{trabelsi_deep_2017}, we use it for automatic music
transcription task and study the $1d$ VGG-like $\cplx$-convent proposed therein. Similarly to
the study, we downsample the audio from 44.1kHz to 11kHz, retain only $84$ out of $128$ labels,
and hold-out the same validation and test splits. During each training epoch we sample $1000$
mini-batches of the remaining $321$ compositions, with each mini-batch comprised of a random
full window of $4096$ samples picked from each waveform and transformed from $\real$ to
$\cplx$ with $1d$ FFT. Annotation labels are taken at the mid-point of the window.

Preliminary experiments with the model without sparsification have shown that early stopping
almost always terminates the training process within the first $10-20$ epochs with the validation
performance peaking at $10-15$ epochs, much sooner than $200$ epochs used by \citet{trabelsi_deep_2017}.
Thus we adopt shorter lasting stages: $12$, $32$ and $50$ epochs, which essentially turn the
``dense'' stage into pretraining. Early stopping is used only during the ``fine-tune'' stage.
To keep the learning rate schedule consistent, we allow the base rate of $10^{-3}$ to be scaled
after $5$, $10$, and $20$-th epoch by $\tfrac1{10}$, $\tfrac1{20}$ and $\tfrac1{100}$, respectively.
Other key deviations from the set-up used by \citet{trabelsi_deep_2017} include: shifting lower
frequencies to the centre of the window to maintain locality after applying the FFT
(sec.~\ref{sub:mnist_like_datasets}), and clipping $\ell_2$ norm of the gradients to $0.05$.

\begin{figure}[!h]
  \centering
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{../assets/figure__musicnet__trade-off/paper__musicnetram__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    Performance-compression trade-off using \textbf{\color{blue} VD} and \textbf{\color{orange} ARD}.
    The \textbf{\color{pink} version} of the network with kernel $3$ in the first convolution is
    compressed with VD only.
  }
  \label{fig:musicnet__trade-off}
\end{figure}

In each experiment we vary the method and the value $C$, while keeping $\tau$ at $-\tfrac12$.
Every experiment is replicated five times to take into account the random effects from
initialization, batch sequence during SGD, and stochasticity of intermediate layers on
forward pass at ``sparsify'' stage.

Figure~\ref{fig:musicnet__trade-off} shows the resulting performance-compression frontier
for the deep network of \citet{trabelsi_deep_2017} and its version, in which we purposefully
halve the spatial size of the first $1d$ convolution kernel from $6$ to $3$ (denoted by
suffix \textit{k}$3$). The motivation is to test if the handicap introduced by the forced
compression of the most upstream layer can be alleviated through non-uniform compression,
induced by variational dropout. We test only VD (sec.~\ref{ssub:variational_dropout}) this
sub-experiment, since prior results have not demonstrated significant superiority of one method
over another (sec.~\ref{ssub:empirical_bayes}).

For compression levels lower than $50$, the final stage overfits the network, since it still
remains overparameterized. For the rates higher than $\approx 100$ the ``fine-tune'' stage
serves its purpose, essentially, due regularizing effects of high sparsity. There is evidence
in favour of achievability of the SOTA average precision of \citet{trabelsi_deep_2017} within
the $50$-$100$ compression region with $\cplx$-valued variational dropout with ARD prior
(sec.~\ref{ssub:empirical_bayes}).

@@

Despite having higher arithmetic complexity, complex-valued models provide built-in
amplitude-phase transformations and regularization \citep{hirose_complex-valued_2009}
and enjoy superior performance on naturally $\cplx$-valued datasets \citep{tarver_design_2019}.
In this study we address the issue of excessive arithmetic complexity of complex-valued
networks by extending the existing Bayesian compression techniques by \citet{kingma_variational_2015},
\citet{molchanov_variational_2017} and \citet{kharitonov_variational_2018} to $\cplx$-valued
parameters. To this end we have conducted a large numerical study of such networks to assess
the performance-compression rate trade-off and have achieved performance comparable to
the state-of-the-art in $\cplx$-valued networks on the MusicNet dataset.

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=\linewidth]{../assets/grad_log.pdf}
  \caption{$\tfrac{d f(\alpha)}{d \log{\alpha}}$ of the fit proposed by
  \citet{molchanov_variational_2017}, MC estimate of \eqref{eq:improper-kl-div-real},
  and the exact derivative using \eqref{eq:real-kl-div-deriv-log}.}
  \label{fig:molchanov-derivative-replica}
\end{figure}
